Refinement of Thermostated Molecular Dynamics via Backward Error Analysis

### 2018-06-14 rigid_bodies.tex
Changes in Discretization Errors section

\setkeys{acs}{
	abbreviations = false,
	articletitle  = false,
	keywords      = false,
	maxauthors    = 10,
	super         = true
}
% Comment below before submitting
\let\titlefont\undefined
\usepackage[fontsize=11pt]{scrextend}
\usepackage[hidelinks,colorlinks,citecolor=blue]{hyperref}
%\flushbottom
% Comment above before submitting

\usepackage{xr}
\externaldocument{rigid_bodies_supp_info}
\usepackage{booktabs}
\usepackage{threeparttable}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[T1]{fontenc}
\usepackage[table]{xcolor}
\definecolor{lightgray}{gray}{0.9}
\usepackage{tikz}
\usetikzlibrary{decorations.text}
\newcommand{\mt}[1]{\boldsymbol{\mathbf{#1}}}   % matrix symbol
\newcommand{\vt}[1]{\boldsymbol{\mathbf{#1}}}   % vector symbol
\newcommand{\tr}[1]{#1^\text{t}}                % transposition
\newcommand{\diff}[2]{\frac{\partial #1}{\partial #2}} % derivative
\newcommand{\sdiff}[2]{\nabla^2_{#2}#1}         % 2nd derivative
\newcommand{\Ham}[1]{{\mathcal H}_\text{#1}}    % Hamiltonian
\newcommand{\Liu}[1]{i\!L_\text{#1}}            % Liouville
\newcommand{\timestep}{h}
\newcommand{\refined}[1]{\widetilde{#1}}

@@

@@

\subsection{Discretization Effects on Equilibrium Thermodynamic Properties (Work in Progress)}

%In the previous section, we have shown that the extended energy drifts in NVT simulations are localized within the thermostat variables.
%It is then reasonable to ask if this leads to a more accurate sampling of the physical system.
%To answer this question, we turn our attention to discretization effects.

In Fig.~\ref{fig:properties}, we show the computed averages of some thermodynamic properties as a function of $\timestep$. The simulations were carried out by specifying $T = 298~\text{K}$ and the analyzed quantities correspond to temperature ($T$), potential energy ($U$) and virial ($W$) per molecule, as well as the constant-volume specific heat capacity ($C_V$).
The lines are a guide for the eyes. 
Also, the horizontal dotted lines (black) help to visualize the departure of the computed averages with increasing $\timestep$.
They correspond to the results obtained employing $\timestep = 1~\text{fs}$ in the Refined NHC numerical scheme.
We have included the results of applying our method, but using the approximate solution of rotation due to \citeauthor{Miller_2002} \cite{Miller_2002}, which is denoted as ``Refined/NO-SQUISH''.
Recall that the computed averages for both refined schemes are obtained via reweighting by employing Eq.~\eqref{eq:reweighting}. 
In addition, it should be pointed out that the temperature values in Fig.~\ref{fig:properties}(a) were computed from the simulated kinetic energies assuming equipartition, i.e. $2 \langle K \rangle = (6N -3) k_b T$.

>>>>

\setkeys{acs}{
	abbreviations = false,
	articletitle  = false,
	keywords      = false,
	maxauthors    = 10,
	super         = true
}

% Comment below before submitting
\let\titlefont\undefined
\usepackage[fontsize=11pt]{scrextend}
\usepackage[hidelinks,colorlinks,citecolor=blue]{hyperref}
%\flushbottom
% Comment above before submitting

\usepackage{xr}
\externaldocument{rigid_bodies_supp_info}
\usepackage{booktabs}
\usepackage{threeparttable}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[T1]{fontenc}
\usepackage[table]{xcolor}
\definecolor{lightgray}{gray}{0.9}
\usepackage{tikz}
\usetikzlibrary{decorations.text}

\newcommand{\mt}[1]{\boldsymbol{\mathbf{#1}}}   % matrix symbol
\newcommand{\vt}[1]{\boldsymbol{\mathbf{#1}}}   % vector symbol
\newcommand{\tr}[1]{#1^\text{t}}                % transposition
\newcommand{\diff}[2]{\frac{\partial #1}{\partial #2}} % derivative
\newcommand{\sdiff}[2]{\nabla^2_{#2}#1}         % 2nd derivative
\newcommand{\Ham}[1]{{\mathcal H}_\text{#1}}    % Hamiltonian
\newcommand{\Liu}[1]{i\!L_\text{#1}}            % Liouville
\newcommand{\timestep}{h}
\newcommand{\refined}[1]{\widetilde{#1}}
\newcommand{\mini}[1]{\text{\tiny{#1}}}

@@

\date{\today}

@@

\subsection{Correction of Discretization Effects on Thermodynamic Properties}

In the previous section, we have seen that drifts in the extended energy of an NVT simulation can accumulate in the thermostat-related variables, thus preventing full-fledged instability from happening.
This allows simulations to be executed with large time steps, which could in principle save computer time.
However, as we will show now, discretization errors can become manifest much earlier than any risk of instability.
Fortunately, our proposed methodology is capable of mitigating the effect of these errors on the ensemble averages computed for the simulated system.

In Fig.~\ref{fig:properties}, we present some thermodynamic properties computed for our liquid water system using several NVT-MD methods with different time step sizes.
All simulations were carried out by specifying $T = 298~\text{K}$ and the analyzed quantities are the estimated temperature $T_\text{est} = \frac{2\langle K \rangle}{(6N-3)k_\mini{B}}$,
the mean potential energy per molecule $\langle U/N \rangle$,
the mean internal virial per molecule $\langle W/N \rangle$, and
the constant-volume specific heat $c_\mini{V} = \frac{\langle \Ham{}^2 \rangle - \langle \Ham{} \rangle^2}{N k_\mini{B} T^2}$.
The internal virial $W$ is computed as described in Ref.~\citenum{Silveira_2017}.
All symbols in Fig.~\ref{fig:properties} represent results obtained after $18~\text{ns}$ of production time, while the lines are just guides for the eyes.
The horizontal, dotted lines are meant to help visualizing the departure of the computed averages from their expected values.
They correspond to the results obtained by employing the Refined NHC method with $\timestep = 1~\text{fs}$.
Error bars were obtained using the Overlapping Batch Means estimator \cite{Meketon_1984, Flegal_2010} and standard uncertainty propagation formulas.

We have included the results of applying our method, but using the approximate solution of rotation due to \citeauthor{Miller_2002} \cite{Miller_2002}, which is denoted as ``Refined/NO-SQUISH''.
Recall that the computed averages for both refined schemes are obtained via reweighting by employing Eq.~\eqref{eq:reweighting}.
